Red, Blue, or Purple? 2024’s Battleground States Tell a Story

### 2024-11-04 paper/paper.qmd
paper wording corrected

---
title: "Red, Blue, or Purple? 2024’s Battleground States Tell a Story"
subtitle: "Pennsylvania, Ohio, and Nevada Become Game Changers as Trump Gains Ground"
author: 
  - Yingke He
  - Ziheng Zhong
thanks: "Code and data are available at: [https://github.com/iJustinn/Election_Prediction.git](https://github.com/iJustinn/Election_Prediction.git)."
date: today
date-format: long
abstract: "This paper analyzes polling data for the 2024 U.S. Presidential election, examining key candidates, notably Donald Trump and Kamala Harris, to gain a comprehensive perspective on the race. A Hierarchical Bayesian Model is applied in this study, and the findings reveal significant variations across states, with Pennsylvania, Ohio, Georgia, and Nevada emerging as particularly influential battlegrounds that ultimately tilt in favor of Trump. By highlighting these trends, this research underscores the dynamic and evolving nature of the U.S. political landscape and the crucial role that swing states play in determining election outcomes."
format: pdf
toc: true
number-sections: true
bibliography: references.bib
---

@@

In the lead-up to the 2024 U.S. Presidential election, understanding voter behavior across states and demographic groups is essential for accurately predicting electoral outcomes. Recent polling data from various sources suggests a close race among major candidates, including Donald Trump, Kamala Harris, and other key contenders. This study applies a Hierarchical Bayesian Model to analyze polling data across all U.S. states, with a focus on identifying critical swing states where even slight shifts in voter sentiment could prove decisive. By exploring trends in voter support across different states and demographic groups, this research aims to highlight the regions most likely to influence the final election outcome.

The primary objective of this study is to estimate state-level vote shares for each candidate while accounting for factors such as regional variations, pollster-specific biases, and the timing of polls relative to Election Day. Predictors including polling percentages, sample sizes, and days remaining until the election are incorporated into the model, which adjusts for state and pollster effects. The model’s structure enables the estimation of vote shares that reflect the underlying distribution of polling data across diverse regions in the U.S.

@@

This research contributes to the field of election forecasting by combining aggregated polling data with a robust modeling approach, offering valuable insights for political analysts, campaign strategists, and policymakers. By identifying critical swing states and accounting for potential polling biases, this study equips stakeholders with the tools to anticipate voter shifts and strategize effectively in an evolving political landscape.

The structure of the paper is organized as follows: following @sec-intro, @sec-data outlines the data collection and cleaning process, along with a description of the outcome and predictor variables used in the analysis.@sec-model, introduces the forecasting models and discuss the rationale behind choosing these models for election outcomes prediction. @sec-result then presents the main findings, including a breakdown of state-level and pollster-level random effects. Finally, @sec-discussion interprets the results, highlighting significant trends and predictions, and concludes with a discussion on the reliability of the forecasts and potential limitations of the models.

@@

## Overview [info tobe update]

Overview text [To be updated...]

@@

In this study, polling data shows the public opinion across different states, measuring support for candidates ahead of the 2024 U.S. Presidential election. The data originates from various polling organizations, each having distinct methodologies that influence data quality and representativeness. Therefore, understanding these methodologies is crucial for interpreting polling data accurately. Below shows the primary categories of measurement methods include:

1. Live Telephone Interviews
Some polling organizations, such as Selzer & Co. and Marist College, conduct live telephone interviews, reaching respondents on both landlines and cell phones. This method often provides more accurate and in-depth data because live interviewers can clarify questions, verify demographic details, and increase the engagement of respondants. However, it costs more and is more time-intensive, leading to smaller sample sizes and fewer surveys compared to other methodologies. In the real world, live telephone interviews include respondants from a cross-section of the population, though they may over-represent older individuals who are more likely to answer landline calls.

2. Interactive Voice Response (IVR)
Pollsters like Trafalgar Group and Rasmussen Reports frequently use IVR, or “robo-polling,” where an automated system asks questions over the phone. IVR is more affordable and efficient for reaching large numbers of people. However, because respondents interact with a machine rather than a live person, they may feel less invested in the survey, impacting response rates and depth of engagement. IVR allows pollsters to obtain timely snapshots of voter intentions, though it may miss out on nuanced responses that a live interviewer could capture.

3. Online Panels
YouGov and Ipsos are known for using online panels, which recruit participants via web platforms to answer surveys electronically. This method allows for a fast data collection from a broad audience and is relatively cost-effective. However, online surveys have disadvantages like self-selection bias, as individuals who are more politically active or internet-savvy are more likely to participate. Other than this, online panels are still popular due to their scalability and cost-efficiency, providing a contemporary approach to measuring public opinions.

4. Mixed-Mode Surveys (Telephone and Online)
To balance the benefits and limitations of each mode, organizations like The New York Times/Siena College and Public Policy Polling employ mixed-mode surveys. By combining online and telephone interviews, these pollsters aim to broaden demographic reach and mitigate biases that are in any single method. Mixed-mode surveys improve representativeness and accommodate diverse respondent preferences, providing data that better reflect the full side of public opinion.

5. Specialized In-Person and University-Affiliated Research
Some university-affiliated pollsters, such as Quinnipiac University and University of New Hampshire Survey Center, incorporate face-to-face interviews or demographic-specific focus groups for deeper insights. These methods can yield highly detailed responses and foster a comprehensive understanding of specific voter groups, though they are resource-intensive and typically used for more specialized research.

These varied data collection methods directly affect how public opinion is measured and interpreted from different pollsters. For instance, live interviews may capture a more detailed view of voter sentiment but often exclude younger demographics who primarily use mobile devices or interact online. Meanwhile, IVR and online panels make large-scale data collection affordable and timely, though they may lack depth and introduce potential self-selection biases. Mixed-mode approaches provide a balanced perspective, catering to both traditional and digital respondents, which may yield more representative data.

The choice of methodology reflects each pollster’s objectives and resources, with implications for accuracy and data interpretation. By categorizing pollsters based on their methodologies, we can contextualize the polling data in this study, recognizing each method's strengths and limitations. This understanding is essential for modeling real-world voter behavior accurately, as it acknowledges how data collection methods impact polling results.

@@

To predict the election vote share for each candidate across states, this study employs a Hierarchical Bayesian Model. This model is designed to capture variations in voter support at the state level while accounting for differences introduced by pollsters and specific poll attributes. By structuring the model hierarchically, it accommodates state-level effects, ensuring that local influences on candidate preference are reflected in the predictions. Additionally, the model incorporates poll-specific factors—such as poll sample size, timing, and reliability scores—to enhance its sensitivity to the varying quality and timing of polling data. This comprehensive approach allows the model to generate robust estimates of expected vote shares, providing a thorough view of likely election outcomes.

@@

```
Shown in @tbl-candidate_fixed_effect, the candidate fixed effects in this model capture the baseline level of support for each candidate, reflecting their overall popularity or favorability across the polls. For example, Kamala Harris has a candidate fixed effect of 47.22, while Donald Trump has a slightly lower effect of 45.28. These values suggest that, on average, Harris may have a marginally higher baseline level of support in the polling data compared to Trump. These fixed effects allow the model to adjust for inherent differences in candidate popularity, providing a more accurate prediction by factoring in each candidate's overall favorability independent of state, pollster, or specific poll characteristics. This adjustment is crucial for capturing the candidates' general standing in the electorate, contributing to more reliable forecasting.

@@

The @fig-presidential-pies-3, illustrates the direct comparison between Trump, Harris, and the "Others" category. In this model, Trump holds a slight lead with 46.11%, while Harris closely follows with 45.65%. Meanwhile, 8.24% of the vote remains attributed to "Others," representing voters who may support third-party candidates or are undecided. This small lead for Trump, amounting to just a 0.46% difference, emphasizes the closeness of the race. The presence of a significant "Others" portion could prove crucial if those votes shift toward one of the primary candidates, potentially altering the predicted outcome.

@@

This pattern underscores the competitive nature of the election, where both candidates have strongholds, but the overall lead may still be within reach for either. The close percentages—often within 1-2% margins—highlight the importance of final campaign efforts and voter turnout in swaying these crucial states.

@@

The inclusion of state effects in the model allows for more accurate predictions that account for both national trends and state-specific deviations. This approach enhances the reliability of the model by recognizing and adjusting for regional variations, leading to predictions that more accurately reflect the complexities of voter sentiment across the United States. By accommodating these state-level effects, the model can better provide insights that are both comprehensive and reflective of the current diverse political dynamics.

@@

The recency adjustment in the model gives more weight to polls conducted closer to the election, which can enhance the model’s responsiveness to recent changes in voter sentiment. This adjustment is crucial if recent events have significantly impacted public opinion, as it allows the model to prioritize recent data. In this case, the plot suggests that while recency adjustments provide added precision by refining the distribution, they do not drastically alter the predictions. This alignment between the original and recency-adjusted predictions may indicate steady voter sentiment, with minimal late shifts influencing the overall distribution as the election date approaches.

@@

The primary data source for this study is the national polling data on the 2024 U.S. Presidential election provided by FiveThirtyEight. This dataset compiles polling results from a variety of polling organizations, each employing distinct methodologies, sample sizes, and timing intervals. By consolidating polls from multiple sources, FiveThirtyEight’s dataset offers a comprehensive view of public opinion trends across the nation, capturing fluctuations in candidate support as the election approaches. This centralized dataset also includes detailed information on each pollster's historical performance and reliability, enabling a more nuanced understanding of poll accuracy and biases.

>>>>

---
title: "Red, Blue, or Purple? 2024’s Battleground States Tell a Story"
subtitle: "Pennsylvania, Ohio, and Nevada Become Game Changers as Trump Gains Ground"
author: 
  - Yingke He
  - Ziheng Zhong
thanks: "Code and data are available at: [https://github.com/iJustinn/Election_Prediction.git](https://github.com/iJustinn/Election_Prediction.git)."
date: today
date-format: long
abstract: "This paper presents an analysis of polling data for the 2024 U.S. Presidential election, focusing on key candidates, notably Donald Trump and Kamala Harris, to provide an in-depth assessment of the race. Employing a Hierarchical Bayesian Model, the study reveals substantial state-level variations, with Pennsylvania, Ohio, Georgia, and Nevada identified as pivotal battlegrounds that show a tendency to lean toward Trump. These findings emphasize the dynamic and shifting landscape of U.S. politics and underscore the critical impact of swing states on election results."
format: pdf
toc: true
number-sections: true
bibliography: references.bib
---

@@

In the approach to the 2024 U.S. Presidential election, analyzing voter behavior across states and demographic groups is vital for improving the accuracy of electoral outcome predictions. Recent polling data indicate a close competition among primary candidates, including Donald Trump and Kamala Harris, underscoring the importance of precise forecasting. This study employs a Hierarchical Bayesian Model to analyze polling data from all U.S. states, with particular attention to identifying swing states where minor shifts in voter sentiment could significantly impact the results. By examining trends in voter support across states and demographics, this research aims to enhance the accuracy of election predictions by pinpointing the regions most likely to influence the final electoral outcome.

The study's main objective is to estimate state-level vote shares for each candidate, taking into account variables such as regional differences, pollster-specific biases, and the timing of polls relative to Election Day. Key predictors—including polling percentages, sample sizes, and days remaining until the election—are incorporated into the model, which adjusts for variations by state and pollster. The Hierarchical Bayesian Model was selected for its capability to capture intricate data patterns, enabling accurate vote share estimations that mirror the distribution of polling data across diverse U.S. regions.

@@

This research contributes to the field of election forecasting by integrating aggregated polling data with a robust modeling approach, offering insights for political analysts, campaign strategists, and policymakers. By identifying pivotal swing states and addressing possible polling biases, this study provides stakeholders with tools to anticipate shifts in voter preferences and adapt strategies effectively in a dynamic political landscape.

The structure of the paper is organized as follows: following @sec-intro, @sec-data outlines the data collection and cleaning process, along with a description of the outcome and predictor variables used in the analysis. @sec-model, introduces the forecasting models and discuss the rationale behind choosing these models for election outcomes prediction. @sec-result then presents the main findings, including a breakdown of state-level and pollster-level random effects. Finally, @sec-discussion interprets the results, highlighting significant trends and predictions, and concludes with a discussion on the reliability of the forecasts and potential limitations of the models.

@@

## Source

The dataset for this analysis is derived from FiveThirtyEight's 2024 National Presidential Polls [@cite538], capturing a broad array of polling data related to the 2024 U.S. Presidential election. This dataset includes survey results from various polling organizations across the country, offering insights into voter preferences by candidate, demographic group, and geographic location. Key variables encompass the polling organization, sample size, polling dates, candidate support percentages, and methodological notes. These elements are essential for interpreting public sentiment over the course of the election campaign, providing snapshots of preferences at distinct points in time. Additionally, variables such as “weight”—applied to adjust for factors like sample size and sample representativeness—enable more precise aggregate predictions. The dataset's context emphasizes its role in informing both the public and analysts about voter behavior trends throughout an election year, underscoring the value of reliable and transparent polling data in building public trust and supporting informed decision-making.

Although alternative datasets, such as those from RealClearPolitics or the Cook Political Report, were available, the FiveThirtyEight dataset was selected due to its detailed methodology, which incorporates adjustments for recent polling data and pollster reliability, thus enabling a more precise analysis of election trends. The dataset underwent necessary preprocessing, which included addressing missing values and standardizing candidate names, as naming conventions and formatting (e.g., use of middle initials) often differ across polling organizations. These inconsistencies were standardized to facilitate accurate comparisons. Additional variables were created, including aggregated candidate support by state and calculated shifts over time, to aid in interpreting broader trends. Summary statistics on each candidate’s average support were calculated, and visualizations captured both national and state-level patterns. Graphs and tables in the analysis depict support levels by demographic and geographic variables, facilitating an in-depth examination of relationships and trends among various voter groups.

@@

In this study, polling data provides insights into public opinion across various states, capturing candidate support leading up to the 2024 U.S. Presidential election. The data is sourced from multiple polling organizations, each employing distinct methodologies that affect data quality and representativeness. Understanding these methodologies is essential for accurately interpreting the polling data. The primary categories of measurement methods include:

1. **Live Telephone Interviews: **
Some polling organizations, such as Selzer & Co. and Marist College, conduct live telephone interviews, reaching respondents on both landlines and cell phones. This method often provides more accurate and in-depth data because live interviewers can clarify questions, verify demographic details, and increase the engagement of respondants. However, it costs more and is more time-intensive, leading to smaller sample sizes and fewer surveys compared to other methodologies. In the real world, live telephone interviews include respondants from a cross-section of the population, though they may over-represent older individuals who are more likely to answer landline calls.

2. **Interactive Voice Response (IVR): **
Pollsters like Trafalgar Group and Rasmussen Reports frequently use IVR, or “robo-polling,” where an automated system asks questions over the phone. IVR is more affordable and efficient for reaching large numbers of people. However, because respondents interact with a machine rather than a live person, they may feel less invested in the survey, impacting response rates and depth of engagement. IVR allows pollsters to obtain timely snapshots of voter intentions, though it may miss out on nuanced responses that a live interviewer could capture.

3. **Online Panels: **
YouGov and Ipsos are known for using online panels, which recruit participants via web platforms to answer surveys electronically. This method allows for a fast data collection from a broad audience and is relatively cost-effective. However, online surveys have disadvantages like self-selection bias, as individuals who are more politically active or internet-savvy are more likely to participate. Other than this, online panels are still popular due to their scalability and cost-efficiency, providing a contemporary approach to measuring public opinions.

4. **Mixed-Mode Surveys (Telephone and Online): **
To balance the benefits and limitations of each mode, organizations like The New York Times/Siena College and Public Policy Polling employ mixed-mode surveys. By combining online and telephone interviews, these pollsters aim to broaden demographic reach and mitigate biases that are in any single method. Mixed-mode surveys improve representativeness and accommodate diverse respondent preferences, providing data that better reflect the full side of public opinion.

5. **Specialized In-Person and University-Affiliated Research: **
Some university-affiliated pollsters, such as Quinnipiac University and the University of New Hampshire Survey Center, use face-to-face interviews or demographic-specific focus groups to gain deeper insights. These methods can provide highly detailed responses and enable a thorough understanding of specific voter segments, although they are resource-intensive and generally reserved for more specialized research purposes.

The diversity of data collection methods directly influences how public opinion is measured and interpreted across pollsters. For example, live interviews often capture nuanced voter sentiment but may underrepresent younger demographics who primarily use mobile devices or engage online. In contrast, IVR and online panel methods enable cost-effective, large-scale data collection but may lack depth and introduce self-selection biases. Mixed-mode approaches attempt to address these limitations by incorporating both traditional and digital respondents, potentially resulting in more representative data.

Each methodology reflects a pollster’s specific objectives and available resources, impacting both the accuracy and interpretability of the data. By categorizing pollsters according to their methodologies, this study seeks to contextualize polling data more effectively, acknowledging the distinct strengths and limitations of each approach. Such awareness is essential for accurately modeling voter behavior, as it highlights the influence of data collection techniques on polling outcomes.

@@

To predict election vote shares for each candidate across states, this study employs a Hierarchical Bayesian Model designed to capture state-level variations in voter support while adjusting for differences introduced by pollsters and specific polling characteristics. The hierarchical structure allows the model to account for state-specific effects, ensuring that regional influences on candidate preferences are accurately represented in the predictions. Additionally, the model incorporates poll-specific factors—such as sample size, timing, and reliability scores—to improve its sensitivity to polling data quality and recency. This approach allows the model to generate robust estimates of expected vote shares, providing a well-rounded perspective on probable election outcomes.

@@

As shown in @tbl-candidate_fixed_effect, the candidate fixed effects in this model represent the baseline support level for each candidate, indicating their overall popularity or favorability across the polling data. For instance, Kamala Harris has a candidate fixed effect of 47.22, while Donald Trump’s effect is slightly lower at 45.28. These values suggest that Harris, on average, may have a marginally higher baseline support level in the polls compared to Trump. By incorporating these fixed effects, the model adjusts for inherent differences in candidate popularity, enabling more accurate predictions by accounting for each candidate's overall favorability, independent of state, pollster, or individual poll characteristics. This adjustment is essential for accurately reflecting each candidate’s general standing within the electorate, thereby enhancing the model’s forecasting reliability.

@@

As illustrated in @fig-presidential-pies-3, the model directly compares support levels for Trump, Harris, and the "Others" category. Trump leads narrowly with 46.11%, followed closely by Harris at 45.65%, while 8.24% of the vote falls into the "Others" category, indicating potential support for third-party candidates or undecided voters. This slight lead for Trump, with a margin of only 0.46%, underscores the race’s tightness. The considerable share of votes in the "Others" category could be pivotal if these voters finally align with one of the main candidates, potentially influencing the final outcome.

@@

This pattern highlights the election’s competitive nature, with each candidate holding strong regional support, yet an overall lead remaining within reach for either. The narrow margins—often within 1-2%—underscore the critical impact of final campaign efforts and voter turnout in influencing outcomes in these pivotal states.

@@

Incorporating state effects into the model allows for more precise predictions that consider both national trends and unique state-level variations. This approach strengthens the model’s reliability by recognizing and adjusting for regional differences, resulting in predictions that more accurately capture the complexities of voter sentiment across the United States. By accounting for these state-specific effects, the model generates insights that are both detailed and reflective of the country’s diverse political landscape.

@@

The recency adjustment in the model increases the weight of polls conducted closer to the election, enhancing the model’s sensitivity to recent shifts in voter sentiment. This adjustment is particularly beneficial when recent events have had a notable impact on public opinion, as it enables the model to prioritize current data. In this case, the plot indicates that while the recency adjustment refines the distribution and improves precision, it does not significantly alter the overall predictions. This consistency between the original and recency-adjusted predictions may suggest stable voter sentiment, with few late changes affecting the overall distribution as the election nears.

@@

The primary data source for this study is national polling data on the 2024 U.S. Presidential election, provided by FiveThirtyEight (@cite538). This dataset compiles polling results from a range of organizations, each employing distinct methodologies, sample sizes, and timing intervals. By integrating data from multiple sources, FiveThirtyEight provides a broad view of national public opinion trends, capturing fluctuations in candidate support as the election approaches. Additionally, the dataset includes detailed information on each pollster's historical performance and reliability, supporting a more nuanced evaluation of poll accuracy and potential biases.
